Geometric signs reveal changes in social structures and networks during the Western Eurasian Aurignacian

### 2025-08-22 paper/paper.qmd
editing
conclusion edits - add in Kovacevic paper

@@

This study adds to our knowledge of social networks within the Aurignacian period, as our results document dynamic patterns of social exchange and connection. Previous work [@bahn1982inter; @Horiuchi_Takakura_2019; @Rogers_2013; @baker2024evidence; @Golovanova_Doronichev_Doronicheva_Sapega_Shackley_2021; @Gamble_1982] has used material culture analysis to reconstruct Upper Paleolithic social networks, which have been argued to be a driver of cultural and evolutionary change [@dErrico_Baker_Rigaud_2025; @Greenbaum_Friesem_Hovers_Feldman_Kolodny_2019; @Creanza_Kolodny_Feldman_2017; @Cullen_1996; @Hill_Wood_Baggio_Hurtado_Boyd_2014; @Derex_Boyd_2016; @Romano_Lozano_Fern2022]. @Foley_Gamble_2009 places the emergence of complex inter-group connectivity at around 30 k BP, however, our results show evidence of inter-connectivity beginning at around 41 k BP, perhaps suggesting an earlier emergence of complex inter-connectivity than previously predicted. We attempted to rectify the low temporal resolution that tends to affect network analysis results when analyzing older, more time-averaged data sets [@Gravel-Miguel_Coward_2023] by dividing the data up into smaller time periods. Additionally, by focusing on change over time, we are able to see social networks as fluid and dynamic, rather than static and bounded [@Maher_Conkey_2019].

>>>>

Our results do not suggest that geometric signs serve as markers of a structured ethnic identity, as found with Aurignacian personal ornaments in @Vanhaeren_d. However, the data used by @Vanhaeren_d was also analyzed by @Kovacevic_Shennan_Vanhaeren_dErrico_Thomas_2015 to find that cultural identity by-descent with modification and isolation-by-distance is a more likely explanation for the distribution of Aurignacian personal ornaments. Our results support the pattern of cultural identity by-descent with modification (although they do not support isolation-by-distance [@tbl-mantel-table]), as well as the patterns of signs and other artwork as a form of communication of cultural in-group and out-group boundaries proposed by @Sauvet_2018 and @baker2024evidence.

@@

This study adds to our knowledge of social networks within the Aurignacian period, as our results document dynamic patterns of social exchange and connection. Previous work [@bahn1982inter; @Horiuchi_Takakura_2019; @Rogers_2013; @baker2024evidence; @Golovanova_Doronichev_Doronicheva_Sapega_Shackley_2021; @Gamble_1982] has used material culture analysis to reconstruct Upper Paleolithic social networks, which have been argued to be a driver of cultural and evolutionary change [@dErrico_Baker_Rigaud_2025; @Greenbaum_Friesem_Hovers_Feldman_Kolodny_2019; @Creanza_Kolodny_Feldman_2017; @Cullen_1996; @Hill_Wood_Baggio_Hurtado_Boyd_2014; @Derex_Boyd_2016; @Romano_Lozano_Fern2022]. @Foley_Gamble_2009 places the emergence of complex inter-group connectivity at around 30 k BP, however, our results show evidence of inter-connectivity beginning at around 41 k BP, perhaps suggesting an earlier emergence of complex inter-connectivity than previously predicted. We attempted to rectify the low temporal resolution that tends to affect network analysis results when analyzing older, more time-averaged data sets [@Gravel-Miguel_Coward_2023] by dividing the data up into smaller time periods. Additionally, by focusing on change over time, we are able to see social networks as fluid and dynamic, rather than static and bounded [@Maher_Conkey_2019].Our results do not support the conclusions drawn in @Vanhaeren_d, where Aurignacian artwork is found to be a symbolic marker of structured ethnic identity, but are supported by @Kovacevic_Shennan_Vanhaeren_dErrico_Thomas_2015, which utilizes the same data as @Vanhaeren_d to conclude that cultural identity by-descent with modification is a more likely explanation for the distribution of Aurigancian ornaments.
